String Constraints with Regex-Counting and String-Length Solved Efficiently

### 2023-10-19 2023-1-19/sections/preliminaries.tex
remove vspace
remove denghang comments

\medskip
\noindent 
\emph{Finite automata.} 
A \emph{(nondeterministic) finite automaton} (NFA)  is a tuple $\NFA=(Q, \Sigma, \delta, I, F)$, where $Q$ is a finite set of states, $\Sigma$ is a finite alphabet, $\delta \subseteq Q \times \Sigma \times Q$ is the transition relation, $I \subseteq Q$ is the set of initial states, and $F \subseteq Q$ is the set of final states. For readability, we write a transition $(q, a, q') \in \delta$ as $q \xrightarrow[\delta]{a} q'$ (or simply $q \xrightarrow{a} q'$). \denghang{Note that one initial state is sufficient. A set of initial states is a technical choice to simplify the defination of union operation in Sec \ref{sec:automaton}.}
The \emph{size} of an NFA $\NFA$, denoted by $|\NFA|$, is defined as the number of transitions of $\NFA$.
%
A \emph{run} of $\NFA$ on a string $w = a_1 \cdots a_n$ is a sequence of transitions $q_0 \xrightarrow{a_1} q_1 \cdots q_{n-1} \xrightarrow{a_n} q_n$ such that $q_0 \in I$. The run is \emph{accepting} if $q_n \in F$.
A string $w$ is accepted by an NFA $\NFA$ if there is an accepting run of $\NFA$ on $w$. In particular, the empty string $\varepsilon$ is accepted by $\NFA$ if $I \cap F \neq \emptyset$. The language of $\NFA$, denoted by $\Lang(\NFA)$, is the set of strings accepted by $\NFA$. 
%
An NFA $\NFA$ is said to be \emph{deterministic} if $I$ is a singleton and, for every $q \in Q$ and $a \in \Sigma$, there is at most one state $q' \in Q$ such that $(q, a, q') \in \delta$. We use DFA to denote deterministic finite automata. 
%

It is well-known that finite automata capture regular languages. 
Moreover, the class of regular languages is closed under union, intersection, concatenation, Kleene star, complement, and language difference \cite{HU79}. \sout{The constructions of NFA corresponding to these language operations are list in \ref{appendix:cefa}}

@@

\begin{theorem}[\cite{SSMH04}]\label{thm-nfa-parikh}
Given an NFA $\NFA$,  an existential LIA formula $\psi_\NFA(\fZ_\Sigma)$ can be computed in linear time that defines the Parikh image of $\NFA$.
\end{theorem}
\vspace{-3mm}

>>>>

\medskip
\noindent 
\emph{Finite automata.} 
A \emph{(nondeterministic) finite automaton} (NFA)  is a tuple $\NFA=(Q, \Sigma, \delta, I, F)$, where $Q$ is a finite set of states, $\Sigma$ is a finite alphabet, $\delta \subseteq Q \times \Sigma \times Q$ is the transition relation, $I \subseteq Q$ is the set of initial states, and $F \subseteq Q$ is the set of final states. For readability, we write a transition $(q, a, q') \in \delta$ as $q \xrightarrow[\delta]{a} q'$ (or simply $q \xrightarrow{a} q'$). Note that one initial state is sufficient. A set of initial states is a technical choice to simplify the defination of union operation in Sec \ref{sec:automaton}.
The \emph{size} of an NFA $\NFA$, denoted by $|\NFA|$, is defined as the number of transitions of $\NFA$.
%
A \emph{run} of $\NFA$ on a string $w = a_1 \cdots a_n$ is a sequence of transitions $q_0 \xrightarrow{a_1} q_1 \cdots q_{n-1} \xrightarrow{a_n} q_n$ such that $q_0 \in I$. The run is \emph{accepting} if $q_n \in F$.
A string $w$ is accepted by an NFA $\NFA$ if there is an accepting run of $\NFA$ on $w$. In particular, the empty string $\varepsilon$ is accepted by $\NFA$ if $I \cap F \neq \emptyset$. The language of $\NFA$, denoted by $\Lang(\NFA)$, is the set of strings accepted by $\NFA$. 
%
An NFA $\NFA$ is said to be \emph{deterministic} if $I$ is a singleton and, for every $q \in Q$ and $a \in \Sigma$, there is at most one state $q' \in Q$ such that $(q, a, q') \in \delta$. We use DFA to denote deterministic finite automata. 
%

It is well-known that finite automata capture regular languages. 
Moreover, the class of regular languages is closed under union, intersection, concatenation, Kleene star, complement, and language difference \cite{HU79}. The constructions of NFA corresponding to these language operations are list in \ref{appendix:cefa}

@@

\begin{theorem}[\cite{SSMH04}]\label{thm-nfa-parikh}
Given an NFA $\NFA$,  an existential LIA formula $\psi_\NFA(\fZ_\Sigma)$ can be computed in linear time that defines the Parikh image of $\NFA$.
\end{theorem}
